### 2015-04-13 datasummary.tex

\documentclass{article} 
\usepackage{amssymb,amsfonts,amsmath}
\usepackage{color,graphicx}
\usepackage[left=0.5cm,right=0.5cm,top=0.5cm,bottom=0.5cm]{geometry}
\newcommand{\picdir}{pdffig}
\begin{document}
\section{Data Summary}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0530/magnbestfit0018.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0530/roimrtibestfit0018.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0530/roisembestfit0018.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0530/roimrtidosebestfit0018.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0530/roisemdosebestfit0018.jpg}} \end{tabular} \caption{Study0035,0530, quality 1, Unknown, obj 84986.645172, dce 0.942222  } \label{Study00350530} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0433/magnbestfit0018.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0433/roimrtibestfit0018.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0433/roisembestfit0018.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0433/roimrtidosebestfit0018.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0433/roisemdosebestfit0018.jpg}} \end{tabular} \caption{Study0023,0433, quality 1, Unknown, obj 197923.759766, dce 0.000000  } \label{Study00230433} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0428/magnbestfit0095.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0428/roimrtibestfit0095.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0428/roisembestfit0095.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0428/roimrtidosebestfit0095.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0428/roisemdosebestfit0095.jpg}} \end{tabular} \caption{Study0023,0428, quality 1, Unknown, obj 363041.350220, dce 0.472527  } \label{Study00230428} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0495/magnbestfit0142.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0495/roimrtibestfit0142.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0495/roisembestfit0142.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0495/roimrtidosebestfit0142.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0495/roisemdosebestfit0142.jpg}} \end{tabular} \caption{Study0030,0495, quality 1, Unknown, obj 1154127.152344, dce 0.734291  } \label{Study00300495} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0497/magnbestfit0043.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0497/roimrtibestfit0043.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0497/roisembestfit0043.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0497/roimrtidosebestfit0043.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0497/roisemdosebestfit0043.jpg}} \end{tabular} \caption{Study0030,0497, quality 0, Unknown, obj 27052.624237, dce 0.000000  } \label{Study00300497} \end{figure}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0488/magnbestfit0060.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0488/roimrtibestfit0060.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0488/roisembestfit0060.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0488/roimrtidosebestfit0060.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0488/roisemdosebestfit0060.jpg}} \end{tabular} \caption{Study0030,0488, quality 1, Unknown, obj 401542.793823, dce 0.867514  } \label{Study00300488} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0491/magnbestfit0083.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0491/roimrtibestfit0083.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0491/roisembestfit0083.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0491/roimrtidosebestfit0083.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0491/roisemdosebestfit0083.jpg}} \end{tabular} \caption{Study0030,0491, quality 0, Unknown, obj 254509.986694, dce 0.710335  } \label{Study00300491} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0496/magnbestfit0076.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0496/roimrtibestfit0076.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0496/roisembestfit0076.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0496/roimrtidosebestfit0076.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0496/roisemdosebestfit0076.jpg}} \end{tabular} \caption{Study0030,0496, quality 0, Unknown, obj 324150.427124, dce 0.867368  } \label{Study00300496} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0490/magnbestfit0040.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0490/roimrtibestfit0040.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0490/roisembestfit0040.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0490/roimrtidosebestfit0040.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0490/roisemdosebestfit0040.jpg}} \end{tabular} \caption{Study0030,0490, quality 0, Unknown, obj 81196.404205, dce 0.772627  } \label{Study00300490} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0378/magnbestfit0077.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0378/roimrtibestfit0077.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0378/roisembestfit0077.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0378/roimrtidosebestfit0077.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0378/roisemdosebestfit0077.jpg}} \end{tabular} \caption{Study0017,0378, quality 0, Unknown, obj 287035.737564, dce 0.787149  } \label{Study00170378} \end{figure}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0402/magnbestfit0146.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0402/roimrtibestfit0146.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0402/roisembestfit0146.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0402/roimrtidosebestfit0146.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0402/roisemdosebestfit0146.jpg}} \end{tabular} \caption{Study0018,0402, quality 1, Unknown, obj 892903.752808, dce 0.850812  } \label{Study00180402} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0389/magnbestfit0084.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0389/roimrtibestfit0084.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0389/roisembestfit0084.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0389/roimrtidosebestfit0084.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0389/roisemdosebestfit0084.jpg}} \end{tabular} \caption{Study0018,0389, quality 1, Unknown, obj 426052.625977, dce 0.860215  } \label{Study00180389} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0385/magnbestfit0094.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0385/roimrtibestfit0094.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0385/roisembestfit0094.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0385/roimrtidosebestfit0094.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0385/roisemdosebestfit0094.jpg}} \end{tabular} \caption{Study0018,0385, quality 1, Unknown, obj 509835.214417, dce 0.392157  } \label{Study00180385} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0476/magnbestfit0026.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0476/roimrtibestfit0026.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0476/roisembestfit0026.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0476/roimrtidosebestfit0026.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0476/roisemdosebestfit0026.jpg}} \end{tabular} \caption{Study0029,0476, quality 0, Unknown, obj 138755.732666, dce nan  } \label{Study00290476} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0477/magnbestfit0038.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0477/roimrtibestfit0038.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0477/roisembestfit0038.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0477/roimrtidosebestfit0038.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0477/roisemdosebestfit0038.jpg}} \end{tabular} \caption{Study0029,0477, quality 1, Unknown, obj 872647.494385, dce 0.734694  } \label{Study00290477} \end{figure}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0438/magnbestfit0088.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0438/roimrtibestfit0088.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0438/roisembestfit0088.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0438/roimrtidosebestfit0088.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0438/roisemdosebestfit0088.jpg}} \end{tabular} \caption{Study0025,0438, quality 1, Unknown, obj 175767.360535, dce 0.805907  } \label{Study00250438} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0435/magnbestfit0063.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0435/roimrtibestfit0063.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0435/roisembestfit0063.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0435/roimrtidosebestfit0063.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0435/roisemdosebestfit0063.jpg}} \end{tabular} \caption{Study0025,0435, quality 0, Unknown, obj 198726.876099, dce 0.722853  } \label{Study00250435} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0440/magnbestfit0030.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0440/roimrtibestfit0030.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0440/roisembestfit0030.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0440/roimrtidosebestfit0030.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0440/roisemdosebestfit0030.jpg}} \end{tabular} \caption{Study0025,0440, quality 0, Unknown, obj 174983.430054, dce 0.539419  } \label{Study00250440} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0436/magnbestfit0039.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0436/roimrtibestfit0039.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0436/roisembestfit0039.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0436/roimrtidosebestfit0039.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0436/roisemdosebestfit0039.jpg}} \end{tabular} \caption{Study0025,0436, quality 1, Unknown, obj 99520.776550, dce 0.514286  } \label{Study00250436} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0466/magnbestfit0068.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0466/roimrtibestfit0068.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0466/roisembestfit0068.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0466/roimrtidosebestfit0068.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0466/roisemdosebestfit0068.jpg}} \end{tabular} \caption{Study0028,0466, quality 0, Unknown, obj 113783.561768, dce 0.000000  } \label{Study00280466} \end{figure}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0468/magnbestfit0051.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0468/roimrtibestfit0051.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0468/roisembestfit0051.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0468/roimrtidosebestfit0051.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0468/roisemdosebestfit0051.jpg}} \end{tabular} \caption{Study0028,0468, quality 0, Unknown, obj 166961.445404, dce 0.000000  } \label{Study00280468} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0471/magnbestfit0047.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0471/roimrtibestfit0047.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0471/roisembestfit0047.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0471/roimrtidosebestfit0047.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0471/roisemdosebestfit0047.jpg}} \end{tabular} \caption{Study0028,0471, quality 0, Unknown, obj 130838.327881, dce 0.736842  } \label{Study00280471} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0725/magnbestfit0077.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0725/roimrtibestfit0077.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0725/roisembestfit0077.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0725/roimrtidosebestfit0077.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0725/roisemdosebestfit0077.jpg}} \end{tabular} \caption{Study0052,0725, quality 1, Unknown, obj 2127872.824219, dce 0.272727  } \label{Study00520725} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0720/magnbestfit0026.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0720/roimrtibestfit0026.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0720/roisembestfit0026.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0720/roimrtidosebestfit0026.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0720/roisemdosebestfit0026.jpg}} \end{tabular} \caption{Study0052,0720, quality 0, Unknown, obj 209749.526611, dce 0.000000  } \label{Study00520720} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0447/magnbestfit0072.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0447/roimrtibestfit0072.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0447/roisembestfit0072.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0447/roimrtidosebestfit0072.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0447/roisemdosebestfit0072.jpg}} \end{tabular} \caption{Study0026,0447, quality 0, Unknown, obj 219686.516235, dce 0.651515  } \label{Study00260447} \end{figure}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0457/magnbestfit0089.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0457/roimrtibestfit0089.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0457/roisembestfit0089.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0457/roimrtidosebestfit0089.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0457/roisemdosebestfit0089.jpg}} \end{tabular} \caption{Study0026,0457, quality 0, Unknown, obj 218718.776672, dce 0.625698  } \label{Study00260457} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0455/magnbestfit0054.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0455/roimrtibestfit0054.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0455/roisembestfit0054.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0455/roimrtidosebestfit0054.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0455/roisemdosebestfit0054.jpg}} \end{tabular} \caption{Study0026,0455, quality 1, Unknown, obj 144700.032837, dce 0.896552  } \label{Study00260455} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0453/magnbestfit0068.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0453/roimrtibestfit0068.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0453/roisembestfit0068.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0453/roimrtidosebestfit0068.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0453/roisemdosebestfit0068.jpg}} \end{tabular} \caption{Study0026,0453, quality 1, Unknown, obj 381617.614136, dce 0.896266  } \label{Study00260453} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0450/magnbestfit0061.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0450/roimrtibestfit0061.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0450/roisembestfit0061.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0450/roimrtidosebestfit0061.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0450/roisemdosebestfit0061.jpg}} \end{tabular} \caption{Study0026,0450, quality 1, Unknown, obj 275258.614502, dce 0.854430  } \label{Study00260450} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0451/magnbestfit0097.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0451/roimrtibestfit0097.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0451/roisembestfit0097.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0451/roimrtidosebestfit0097.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0451/roisemdosebestfit0097.jpg}} \end{tabular} \caption{Study0026,0451, quality 0, Unknown, obj 234578.714294, dce 0.763251  } \label{Study00260451} \end{figure}
\clearpage
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0418/magnbestfit0086.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0418/roimrtibestfit0086.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0418/roisembestfit0086.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0418/roimrtidosebestfit0086.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0418/roisemdosebestfit0086.jpg}} \end{tabular} \caption{Study0022,0418, quality 1, Unknown, obj 672972.682495, dce 0.845735  } \label{Study00220418} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0417/magnbestfit0027.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0417/roimrtibestfit0027.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0417/roisembestfit0027.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0417/roimrtidosebestfit0027.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0417/roisemdosebestfit0027.jpg}} \end{tabular} \caption{Study0022,0417, quality 0, Unknown, obj 467355.829468, dce 0.500000  } \label{Study00220417} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0409/magnbestfit0028.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0409/roimrtibestfit0028.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0409/roisembestfit0028.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0409/roimrtidosebestfit0028.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0409/roisemdosebestfit0028.jpg}} \end{tabular} \caption{Study0021,0409, quality 0, Unknown, obj 21709.854874, dce nan  } \label{Study00210409} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0414/magnbestfit0094.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0414/roimrtibestfit0094.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0414/roisembestfit0094.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0414/roimrtidosebestfit0094.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0414/roisemdosebestfit0094.jpg}} \end{tabular} \caption{Study0021,0414, quality 1, Unknown, obj 202879.718201, dce 0.000000  } \label{Study00210414} \end{figure}
\begin{figure} \begin{tabular}{ccccc} \scalebox{0.21}{\includegraphics*{\picdir/dakota/0415/magnbestfit0015.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0415/roimrtibestfit0015.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0415/roisembestfit0015.jpg}}  & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0415/roimrtidosebestfit0015.jpg}} & \scalebox{0.21}{\includegraphics*{\picdir/dakota/0415/roisemdosebestfit0015.jpg}} \end{tabular} \caption{Study0021,0415, quality 0, Unknown, obj 27081.701508, dce 0.779412  } \label{Study00210415} \end{figure}
\clearpage
\begin{tabular}{ccc}
\scalebox{0.22}{\includegraphics*{\picdir/UID000}}  
&
\scalebox{0.22}{\includegraphics*{\picdir/UID001}}  
&
\scalebox{0.22}{\includegraphics*{\picdir/UID002}}  
\\
\scalebox{0.22}{\includegraphics*{\picdir/UID003}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID006}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID007}}
\\
\scalebox{0.22}{\includegraphics*{\picdir/UID009}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID010}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID013}}
\\
\scalebox{0.22}{\includegraphics*{\picdir/UID016}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID017}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID018}}
\\
\scalebox{0.22}{\includegraphics*{\picdir/UID019}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID021}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID022}}
\\
\scalebox{0.22}{\includegraphics*{\picdir/UID023}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID024}}
&
\scalebox{0.22}{\includegraphics*{\picdir/UID025}}
\\
\scalebox{0.22}{\includegraphics*{\picdir/UID026}}
\end{tabular}

\end{document}